---
title: "Final Project 2023"
output: html_document
date: "2023-04-23"
---

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(readxl)# use to import .xlsx files
library(ggplot2)
library(ggrepel)
library(GGally)
library(leaflet)
library(sf)
library(RColorBrewer)
library(reshape2)
```


#### Load the dataset, covid-regions.csv dataset.

```{r warning=FALSE, message=FALSE}

regions<-read_csv("./covid-regions.csv") 

head(regions)

```


#### Problem 1

Each group member should create one meaningful question and answer those questions using any descriptive data analysis methods.(Ex:If your group has 4 members, you should have 4 questions)

---

##### Name of the group member who completed this task: Paul Huffman

What is the relation between age, gender, and deaths?
```{r}

# Read the csv file
covid_data <- read.csv("covid-age.csv")

covid_melt <- melt(covid_data, id.vars = c("age_classes"), 
                   measure.vars = c("male_deaths", "female_deaths"), 
                   variable.name = "gender", 
                   value.name = "deaths")

ggplot(covid_melt, aes(x = age_classes, y = deaths, fill = gender)) + 
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Relation between Gender, Age and Deaths",
       x = "Age Classes", y = "Number of Deaths", fill = "Gender") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "top") + scale_fill_discrete(labels=c("Male Deaths", "Female Deaths"))


```

From the above graph it is clear that as age increases so does the number of deaths. It also looks like men are more likely to die from covid as women. The only place where that is not true is in the >90 age group, but I think that is because there are so few people in that age group that it is not a good representation of the population. Also there are more deaths for women than men in that same group, but that is because there are significantly more women who live past 90 than men. Acording to [census.gov](https://www.census.gov/newsroom/releases/archives/aging_population/cb11-194.html) women over 90 outnumber men over 90 at a rate of about 3 to 1.

---

##### Name of the group member who completed this task: Ryan Heiert

```{r}

```

---

##### Name of the group member who completed this task: Justin Bennet

What are the 5 descriptive statistics of the total deaths from covid?
```{r}

boxplot(regions$deaths,
main = "Mean Deaths for Covid in Italy",
xlab = "Total Deaths",
ylab = "Death Count",
col = "orange",
border = "brown",
horizontal = TRUE,
notch = TRUE
)

```

---

##### Name of the group member who completed this task: Christian Lane


```{r}

```

---

#### Problem 2

Use "Circles" option in leaflet package to visualize the Total number of cases. Label the circles by region (Check Module 10 geospatial data visualization)

```{r,warning=FALSE, message=FALSE}

n2<-leaflet(regions) %>% addTiles() %>% addCircles(lng = ~long, lat = ~lat, weight = 1,
    radius = ~total_cases, popup = ~region)

n2

```



#### Problem 3 (Part1) 

Create a parallel plot as given in the project instruction.

Hint: you can use ggplot, GGally, and ggrepel packages 

```{r, warning=FALSE}
ggparcoord(data=regions,
           columns=8:19,
           groupColumn="region") +
  theme(axis.text.x=element_text(angle=90),axis.text = element_text(size = 8)) +
  geom_point(size=NA)+geom_text_repel(aes(label=`...1`))
```

#### Problem 3 (Part2)

Using above plot identify any regions(four regions) which shows deviation from the rest of the regions. Extract the information about those regions. (create a new dataframe, you will use it in next question)


```{r warning=FALSE}

# for some reason it would not allow a collection of length 4, I had to separate like this

newRegions = regions %>% filter(`region` == c("Piemonte", "Lombardia"))
newRegions2 = regions %>% filter(`region` == c( "Veneto", "Emilia-Romagna"))

newRegions = bind_rows(newRegions, newRegions2)

```


#### Problem 3 (Part3)

Use appropriate markers in the leaflet package to visualize the above identified regions on a map. Label the markers with the region name.


```{r}

n3<-leaflet(newRegions) %>% addTiles() %>% addMarkers(~long, ~lat,  label = ~region)

n3

```


#### Problem 4(Part1)

Using mutate function in tidyverse create new data column as  $\mbox{death_per=(deaths/total_cases)*100}$ for each region and call it as "death_per". Add death_per data to the regions dataset. Use mutate function in tidyverse.

```{r}


regions_death_per<-mutate(regions, death_per=(deaths/total_cases)*100)

summary(regions_death_per)

regions<-mutate(regions, death_per=(deaths/total_cases)*100)


```


#### Problem 4 (Part2)

In this problem, your goal is to create a Choropleth map to visualize the Death Percentage for each region. The first steps of the data processing is given. Use the Italy_regions dataframe and complete the code. 

Hint: Module 10 geospatial visualization

```{r, warning=FALSE, message=FALSE}

Italy_regions <- read_sf('shapefile')

name_1=regions$region
regions$name_1<-name_1

 
 # combining data and shapfile 
 is.element(regions$name_1, Italy_regions$name_1) %>%
  all()

 Italy_regions <- merge(Italy_regions, regions, by = 'name_1', all.x = F)
 head(Italy_regions)
 
```


```{r, warning=FALSE, message=FALSE}

paletteNum <- colorNumeric('Blues', domain = Italy_regions$death_per)

leaflet() %>%
  addProviderTiles(providers$CartoDB.PositronNoLabels) %>%
  addPolygons(data = Italy_regions,
              fillColor = ~paletteNum(death_per),
              fillOpacity = 0.7,
              smoothFactor = .3,
              color = "#BDBDC3",
              weight = 1) %>%
  addLegend(pal = paletteNum,
            values = Italy_regions$death_per,
            title = "2020 Percentage  of Deaths<br> by Region in Italy (Covid-19)",
            opacity = 0.7,
            position = 'bottomleft')

```










