---
title: "Final Project 2023"
output: html_document
date: "2023-04-23"
---

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(readxl)# use to import .xlsx files
library(ggplot2)
library(ggrepel)
library(GGally)
library(leaflet)
library(sf)
library(RColorBrewer)
```


#### Load the dataset, covid-regions.csv dataset.

```{r warning=FALSE, message=FALSE}

regions<-read_csv("./covid-regions.csv") 

head(regions)

```


#### Problem 1

Each group member should create one meaningful question and answer those questions using any descriptive data analysis methods.(Ex:If your group has 4 members, you should have 4 questions)

##### Name of the group member who completed this task:..............

```{r}

```


##### Name of the group member who completed this task:..............

```{r}

```


##### Name of the group member who completed this task:..............
What are the 5 descriptive statistics of the total deaths from covid?

```{r}

boxplot(regions$deaths,
main = "Mean Deaths for Covid in Italy",
xlab = "Total Deaths",
ylab = "Death Count",
col = "orange",
border = "brown",
horizontal = TRUE,
notch = TRUE
)

```


##### Name of the group member who completed this task:..............

```{r}

```

#### Problem 2

Use "Circles" option in leaflet package to visualize the Total number of cases. Label the circles by region (Check Module 10 geospatial data visualization)

```{r,warning=FALSE, message=FALSE}

n2<-leaflet(regions) %>% addTiles() %>% addCircles(lng = ~long, lat = ~lat, weight = 1,
    radius = ~total_cases, popup = ~region)

n2

```



#### Problem 3 (Part1) 

Create a parallel plot as given in the project instruction.

Hint: you can use ggplot, GGally, and ggrepel packages 

```{r, warning=FALSE}


```

#### Problem 3 (Part2)

Using above plot identify any regions(four regions) which shows deviation from the rest of the regions. Extract the information about those regions. (create a new dataframe, you will use it in next question)


```{r}

```


#### Problem 3 (Part3)

Use appropriate markers in the leaflet package to visualize the above identified regions on a map. Label the markers with the region name.


```{r}


```


#### Problem 4(Part1)

Using mutate function in tidyverse create new data column as  $\mbox{death_per=(deaths/total_cases)*100}$ for each region and call it as "death_per". Add death_per data to the regions dataset. Use mutate function in tidyverse.

```{r}

```


#### Problem 4 (Part2)

In this problem, your goal is to create a Choropleth map to visualize the Death Percentage for each region. The first steps of the data processing is given. Use the Italy_regions dataframe and complete the code. 

Hint: Module 10 geospatial visualization

```{r, warning=FALSE, message=FALSE}

Italy_regions <- read_sf('shapefile')

name_1=regions$region
regions$name_1<-name_1

 
 # combining data and shapfile 
 is.element(regions$name_1, Italy_regions$name_1) %>%
  all()

 Italy_regions <- merge(Italy_regions, regions, by = 'name_1', all.x = F)
 head(Italy_regions)
 
```


```{r, warning=FALSE, message=FALSE}

```










